### 2016-09-07 notions.tex
cheers

First, a set~$S$ is chosen according to some
distribution~$\distr{}{}$ over the collection of multisets
$\mathcal{S}$ associated with~$\Pi$, and then
$\Rep(S)$ is run to generate $(\pubaux,\privaux)$. \jnote{I thought we were going to let
the adversary choose~$S$?} \tsnote{Email discussion.} \jnote{If instead we choose $S$ from ${\cal D}$, then
is it clear we want to give $S$ to~$A$?} \tsnote{For \emph{correctness},
yes, if for no other reason than to allow~$A$ to know when it asking
a pointless query.  But also it will make no difference to the
provable correctness of the constructions we consider.}
%(Unless stated otherwise, we assume security holds for all distributions and thus for all sets $S$.)
Adversary~$A$ is given~$\pubaux$~as input; it is also
provided with access to an oracle~$\TestOracle$ that, on input a query~$q \in \calQ$,
returns $\Qry(\pubaux,\privaux,q)$ and increments
a counter if this result is in error.  The adversary wins if it
manages to cause at least~$r>0$ such errors among its queries.
\jnote{If we are specializing the set-multiplicity data structures then we
may as well have the attacker query $x$ which gets translated to query $q_x$. If
we are trying to be more general then we should be consistent.}
\tsnote{I don't see how this correctness notion is specialized to
  set-multiplicity data structures.  I don't think we actually
  specialize notions in any way until, maybe, the OW privacy notion.
  It is true that we only deliver concrete results for set-multiplicty
data structures, however.}\jnote{Why does $A$ output
anything at all?}\tsnote{To signify it is finished execution.  Matter
of taste, I'm not hardover about it.}

@@

In this experiment we track the time complexity (relative to some
model of computation) and query complexity (i.e., number of
queries to $\TestOracle$) of~$A$.
We define the advantage of adversary~$A$ in the correctness experiment as
$\AdvCorrect{\Pi,\distr{}{},r}{A} = \Prob{\ExpCorrect{\Pi,\distr{}{}, r}{A}=1}$.
Overloading notation, we write $\AdvCorrect{\Pi,\distr{}{},r}{t,q}$ for the maximum over
all~$t$-time adversaries that ask at most~$q$ queries. We say that a
set-representation data structure~$\Pi$ is $(t,q,r,\epsilon)$-correct if $\AdvCorrect{\Pi,\distr{}{},r}{t,q} \leq \epsilon$.
\jnote{Note~$q$ is overloaded here.}
%\jnote{Note that test includes err because we don't want to penalize
%the adversary's resources for computing $q(S)$.}

\jnote{Case of $r=1$ does not necessarily imply anything for $r>1$: can have a scheme
where finding the first false positive is hard but then the rest are easy; can have
a scheme where finding one false positive is easy finding more is hard (e.g.,
because there do not exist
any more).}

@@

\jnote{Just noticed something (else) odd about the NY definition: we cannot assume w.l.o.g.\
that $A$ makes exactly $q$ queries, and in fact it is possible to have cases where increasing
the number of queries the attacker makes can decrease its advantage! This seems like
another drawback of the definition.}

\begin{figure}[htp]
\centering
\fpage{.6}{
\hpagess{.5}{.45}
{
\experimentv{$\ExpCorrect{\setprim,\distr{}{}}{A}$}\\
$S \getsr \distr{}{}$\\
$\mathcal{Q} \gets \emptyset$\\
%$\mathrm{err}\gets 0$\\
$(\pubaux,\privaux) \getsr \Rep(S)$\\
$q \getsr A^{\TestOracle}(S,\pubaux)$\\
if $\left(q \not \in \mathcal{Q}\right.$ and \\
\nudge $\left.\Qry(\pubaux,\privaux,q) \neq q(S)\right)$ \\
\nudge \nudge then Return 1\\
Return 0
}
%
{
\oracle{$\TestOracle(q)$}\\
$a \gets \Qry(\pubaux,\privaux,q)$\\
$\mathcal{Q} \gets \mathcal{Q} \cup \{q\}$ \\
%\nudge $\mathrm{err}\gets\mathrm{err}+1$\\
Return~$(a,q(S))$
}
}
\caption{The Naor-Yogev definition of
correctness, adapted to our setting.}
\label{fig:NYcorrectness}
\end{figure}

@@

\jnote{Do we have a compelling example of why the definition in Figure~\ref{fig:correctness}
is ``better'' than the definition in Figure~\ref{fig:NYcorrectness}?}

\begin{theorem}
If $\Pi$ is $(t,q,\epsilon)$-NY-correct, then it is also
$(t, q', r, q'\epsilon/r)$-correct for any $q' \leq q+1$ and $r\geq 1$.
\end{theorem}
\begin{proof}
Let $\Pi$ be a set-representation data structure that is $(t,q,\epsilon)$-NY-correct,
and assume that for some distribution $\distr{}{}$, some $q' \leq q+1$, and
some $r \geq 1$
there is an adversary $A$ running in time~$t$
and making $q'$ oracle queries such that
$\AdvCorrect{\Pi,\distr{}{},r}{A} > q'\epsilon/r$.
(Note that
we may assume $A$ always makes exactly $q'$ queries without loss of generality.)
With probability at least
$q'\epsilon/r$ in an execution of $\ExpCorrect{\Pi,\distr{}{},r}{A}$, we have that
$A$ makes at least $r$ distinct queries to $\TestOracle$ for which
an incorrect answer is returned. Let $A'$ be the algorithm that simply
runs~$A$, but chooses
uniformly one of the $q'$ queries of $A$ to its $\TestOracle$ oracle and outputs that query
as its final output. Then with probability at least $r/q' \cdot (q'\epsilon/r)=\epsilon$
the query chosen by $A'$
leads to an incorrect answer, and was not previously asked to the $\TestOracle$ oracle.
Since the running time of $A'$ is at most $t$, and it makes at most $q'-1 \leq q$ queries
to its oracle, this is a contradiction.
\end{proof}

@@

\begin{theorem}
If $\Pi$ is $(t, q, 1, \epsilon)$-correct, then it is also $(t, q-1, \epsilon)$-NY-correct.
\end{theorem}
\begin{proof}
Let $\Pi$ be a set-representation data structure that is $(t,q,1,\epsilon)$-correct,
and assume that for some distribution $\distr{}{}$
there is an adversary $A$ running in time~$t$
and making at most $q-1$ oracle queries such that
$\NYAdvCorrect{\Pi,\distr{}{}}{A} > \epsilon$.
Let $A'$ be the algorithm that simply
runs~$A$, passing the oracle queries of $A$ to its own oracle, until $A$ terminates
with output~$q$; then, $A'$ sends $q$ to $\TestOracle$. \jnote{$q$ is (still)
overloaded!} It is immediate that $A'$ makes at most $q$ oracle queries, and
$\AdvCorrect{\Pi,\distr{}{},1}{A'} \geq \NYAdvCorrect{\Pi,\distr{}{}}{A}$, a contradiction.
\end{proof}

@@

\jnote{I'm not sure what to conclude from any of the results above. The proofs
are also all trivial.}

@@

\jnote{OK, from here on I specialize to set-representation data structures\ldots}

>>>>

First, a set~$S$ is chosen according to some
distribution~$\distr{}{}$ over the collection of multisets
$\mathcal{S}$ associated with~$\Pi$, and then
$\Rep(S)$ is run to generate $(\pubaux,\privaux)$. \jnote{I thought we were going to let
the adversary choose~$S$?} \tsnote{Email discussion. At this point,
I'm happy just making~$S$ part of the experiment ``name'', the
way~$\distr{}{}$ is now.  The distribution seems not to matter for
correctness, at least for the schemes we consdier.} \jnote{If instead we choose $S$ from ${\cal D}$, then
is it clear we want to give $S$ to~$A$?} \tsnote{For \emph{correctness},
yes, if for no other reason than to allow~$A$ to know when it asking
a pointless query.  But also it will make no difference to the
provable correctness of the constructions we consider.}
%(Unless stated otherwise, we assume security holds for all distributions and thus for all sets $S$.)
Adversary~$A$ is given~$\pubaux$~as input; it is also
provided with access to an oracle~$\TestOracle$ that, on input a query~$q \in \calQ$,
returns $\Qry(\pubaux,\privaux,q)$ and increments
a counter if this result is in error.  The adversary wins if it
manages to cause at least~$r>0$ such errors among its queries.
\jnote{If we are specializing the set-multiplicity data structures then we
may as well have the attacker query $x$ which gets translated to query $q_x$. If
we are trying to be more general then we should be consistent.}
\tsnote{I don't see how this correctness notion is specialized to
  set-multiplicity data structures.  I don't think we actually
  specialize notions in any way until, maybe, the OW privacy notion.
  It is true that we only deliver concrete results for set-multiplicty
data structures, however.}\jnote{Why does $A$ output
anything at all?}\tsnote{To signify it is finished execution.  Matter
of taste, I'm not hardover about it.}

@@

In this experiment we track the time complexity (relative to some
model of computation) and query complexity (i.e., number of
queries to $\TestOracle$) of~$A$.
We define the advantage of adversary~$A$ in the correctness experiment as
$\AdvCorrect{\Pi,\distr{}{},r}{A} = \Prob{\ExpCorrect{\Pi,\distr{}{}, r}{A}=1}$.
Overloading notation, we write $\AdvCorrect{\Pi,\distr{}{},r}{t,q}$ for the maximum over
all~$t$-time adversaries that ask at most~$q$ queries. We say that a
set-representation data structure~$\Pi$ is $(t,q,r,\epsilon)$-correct if $\AdvCorrect{\Pi,\distr{}{},r}{t,q} \leq \epsilon$.
\jnote{Note~$q$ is overloaded here.}\tsnote{True, and do we even need
  it, other than for making a compact statements of comparison with
  NY?  Why not just have explict advantage bounding statements, as we
  do for our other results, relating our notion to NY's?  (And,
  probably, moving all of that to an appendix, as you later suggest.)}
%\jnote{Note that test includes err because we don't want to penalize
%the adversary's resources for computing $q(S)$.}

\todo{Explain that the case of $r=1$ does not necessarily imply anything for $r>1$: can have a scheme
where finding the first false positive is hard but then the rest are easy; can have
a scheme where finding one false positive is easy finding more is hard (e.g.,
because there do not exist
any more).}

@@

\jnote{Just noticed something (else) odd about the NY definition: we cannot assume w.l.o.g.\
that $A$ makes exactly $q$ queries, and in fact it is possible to have cases where increasing
the number of queries the attacker makes can decrease its advantage! This seems like
another drawback of the definition.}\tsnote{Really?  That seems worth
pointing out, as part of our list of complaints.}

\begin{figure}[htp]
\centering
\fpage{.6}{
\hpagess{.5}{.45}
{
\experimentv{$\NYExpCorrect{\setprim,\distr{}{}}{A}$}\\
$S \getsr \distr{}{}$\\
$\mathcal{C} \gets \emptyset$\\
%$\mathrm{err}\gets 0$\\
$(\pubaux,\privaux) \getsr \Rep(S)$\\
$q \getsr A^{\TestOracle}(S,\pubaux)$\\
if $\left(q \not \in \mathcal{Q}\right.$ and \\
\nudge $\left.\Qry(\pubaux,\privaux,q) \neq q(S)\right)$ \\
\nudge \nudge then Return 1\\
Return 0
}
%
{
\oracle{$\TestOracle(q)$}\\
$a \gets \Qry(\pubaux,\privaux,q)$\\
$\mathcal{C} \gets \mathcal{C} \cup \{q\}$ \\
%\nudge $\mathrm{err}\gets\mathrm{err}+1$\\
Return~$(a,q(S))$
}
}
\caption{The Naor-Yogev definition of
correctness, adapted to our setting.}
\label{fig:NYcorrectness}
\end{figure}

@@

\jnote{Do we have a compelling example of why the definition in Figure~\ref{fig:correctness}
is ``better'' than the definition in
Figure~\ref{fig:NYcorrectness}?}\tsnote{I don't know, but we do seem
to have a compelling list of shortcomings of their notion, which ours
doesn't seem to suffer. Anyway, it's a bit of an unfair comparison, in
their favor.  You're lifting their notion to ours, which is more
general.  And by the time we get to that, we have already done them
the favor of cleaning up the syntax, so that their ``lifted'' notion
looks better than it was.}

\begin{theorem}
If $\Pi$ is $(t,q,\epsilon)$-NY-correct, then it is also
$(t, q', r, q'\epsilon/r)$-correct for any $q' \leq q+1$ and $r\geq
1$.
\tsnote{this could easily be written, albeit less compactly, without
introducing the resource-parameterized notation; likewise the others.}
\end{theorem}
\begin{proof}
Let $\Pi$ be a set-representation data structure that is $(t,q,\epsilon)$-NY-correct,
and assume that for some distribution $\distr{}{}$, some $q' \leq q+1$, and
some $r \geq 1$
there is an adversary $A$ running in time~$t$
and making $q'$ oracle queries such that
$\AdvCorrect{\Pi,\distr{}{},r}{A} > q'\epsilon/r$.
(Note that
we may assume $A$ always makes exactly $q'$ queries without loss of generality.)
With probability at least
$q'\epsilon/r$ in an execution of $\ExpCorrect{\Pi,\distr{}{},r}{A}$, we have that
$A$ makes at least $r$ distinct queries to $\TestOracle$ for which
an incorrect answer is returned. Let $A'$ be the algorithm that simply
runs~$A$, but chooses
uniformly one of the $q'$ queries of $A$ to its $\TestOracle$ oracle and outputs that query
as its final output. Then with probability at least $r/q' \cdot (q'\epsilon/r)=\epsilon$
the query chosen by $A'$
leads to an incorrect answer, and was not previously asked to the $\TestOracle$ oracle.
Since the running time of $A'$ is at most $t$, and it makes at most $q'-1 \leq q$ queries
to its oracle, this is a contradiction.
\end{proof}

@@

\begin{theorem}
If $\Pi$ is $(t, q, 1, \epsilon)$-correct, then it is also $(t, q-1, \epsilon)$-NY-correct.
\end{theorem}

\begin{proof}
Let $\Pi$ be a set-representation data structure that is $(t,q,1,\epsilon)$-correct,
and assume that for some distribution $\distr{}{}$
there is an adversary $A$ running in time~$t$
and making at most $q-1$ oracle queries such that
$\NYAdvCorrect{\Pi,\distr{}{}}{A} > \epsilon$.
Let $A'$ be the algorithm that simply
runs~$A$, passing the oracle queries of $A$ to its own oracle, until $A$ terminates
with output~$q$; then, $A'$ sends $q$ to $\TestOracle$. \jnote{$q$ is (still)
overloaded!} It is immediate that $A'$ makes at most $q$ oracle queries, and
$\AdvCorrect{\Pi,\distr{}{},1}{A'} \geq \NYAdvCorrect{\Pi,\distr{}{}}{A}$, a contradiction.
\end{proof}

@@

\jnote{I'm not sure what to conclude from any of the results above. The proofs
are also all trivial.} \tsnote{Trivial or not, it's all part of the
overall contribution, which is putting this topic on a solid
foundation.  We can move it all to an appendix, but I still see it as
very worthwhile.}

@@

\jnote{OK, from here on I specialize to set-representation data
  structures\ldots} \tsnote{I don't understand this comment.  That's
  what our syntax supports anyway, so how is this specializing?}

### 2016-09-07 results.tex
starting to typeset ss proof for domain-separated PRF construction
cheers

\begin{theorem}\label{thm1}
Fix $k,m,n,r>0$, and let $\Pi_{\mathrm{lin}}= (\calQ,\Rep, \Qry)$ be the set-multiplicity data structure described in Figure~\ref{fig:lin-and-ds}. Let $\distr{}{}$ be a distribution over~$\mathcal{S}$.  For any adversary~$A$ that makes a total of~$q$ queries to its oracle and has time-complexity~$O(t)$, there exist prf-adversaries~$B_1,B_2$ (explicitly constructed in the proof of this theorem) such that
\[
\AdvCorrect{\Pi_{\mathrm{lin}},\distr{\calS}{},r}{A} \leq  \AdvPRF{F}{B_1} + \AdvPRF{F}{B_2}  +{\dbinom{q}{r}} \left( (1-e^{-kn/m})^k + O(1/n) \right)^r\,.
\]
Here, $B_1$ and $B_2$, each ask~$q$ oracle queries and have time complexity~$O(t+qm)$.
\end{theorem}
Before giving the proof, we note that Kirsch and Mitzenmacher~\cite{xxx} show
that the ($r=1$, zero-query) FP-probability for~$\Pi_\mathrm{lin}$ is
always within $O(1/n)$ of $(1-e^{-kn/m})^k$, although the asymptotic
rate of convergence is faster than~$1/n$.  The factor of $\dbinom{q}{r}$ arises as a result
of moving from the adaptive to the zero-query setting.  Finally, $\dbinom{q}{r} \leq q^r$ (with
reasonable tightness when~$r$ is small) in which case the final term in the bound behaves as
$(q (1-e^{-kn/m})^k + O(q/n) )^r$; loosely, the bound one expects for
trying to find a single FP in each of~$r$ independent``rounds'', each round
consisting of~$q$ attempts.

\heading{Construction of PRF with domain separation.}
Let $\Pi_{\mathrm{ds}}= (\calQ,\Rep, \Qry)$ be a set-multiplicity data structure defined on the right-side of Figure~\ref{fig:lin-and-ds}.  As with the preceding theorem, the following says (informally) that if~$F$ is a secure PRF, then this domain-separated PRF construction is correct against adaptive adversaries.

\begin{theorem}\label{thm2}
Fix $k,m,n,r>0$, and let $\Pi_{\mathrm{ds}}= (\calQ,\Rep, \Qry)$ be the set-multiplicity data structure in Figure~\ref{fig:lin-and-ds}. Let $\distr{}{}$ be a distribution over~$\mathcal{S}$.  For any adversary~$A$ that makes a total of~$q$ queries to its oracle and has time-complexity~$O(t)$, there exists a prf-adversary ~$B$ (explicitly constructed in the proof of this theorem) such that
\[
\AdvCorrect{\Pi_{\mathrm{ds}},\distr{\calS}{},r}{A} \leq  \AdvPRF{F}{B}  + {\dbinom{q}{r}} \left( (1-e^{-kn/m})^k + O(1/n) \right)^r\,.
\]
Here, $B$ asks $q$ queries, and has time complexity $O(t+qm)$.
\end{theorem}

>>>>

\begin{theorem}\label{thm1}
Fix $k,m,n,r>0$, and let $\Pi_{\mathrm{lin}}= (\calQ,\Rep, \Qry)$ be the set-multiplicity data structure described in Figure~\ref{fig:lin-and-ds}. Let $\distr{}{}$ be a distribution over~$\mathcal{S}$.  For any adversary~$A$ that makes a total of~$q$ queries to its oracle and has time-complexity~$O(t)$, there exist prf-adversaries~$B_1,B_2$ (explicitly constructed in the proof of this theorem) such that
\[
\AdvCorrect{\Pi_{\mathrm{lin}},\distr{}{},r}{A} \leq  \AdvPRF{F}{B_1} + \AdvPRF{F}{B_2}  +{\dbinom{q}{r}} \left( (1-e^{-kn/m})^k + O(1/n) \right)^r\,.
\]
Here, $B_1$ and $B_2$, each ask~$q$ oracle queries and have time complexity~$O(t+qm)$.
\end{theorem}
Before giving the proof, we note that Kirsch and Mitzenmacher~\cite{xxx} show
that the ($r=1$, zero-query) FP-probability for~$\Pi_\mathrm{lin}$ is
always within $O(1/n)$ of $(1-e^{-kn/m})^k$, although the asymptotic
rate of convergence is faster than~$1/n$.  The factor of $\dbinom{q}{r}$ arises as a result
of moving from the adaptive to the zero-query setting.  Finally, $\dbinom{q}{r} \leq q^r$ (with
reasonable tightness when~$r$ is small) in which case the final term in the bound behaves as
$(q (1-e^{-kn/m})^k + O(q/n) )^r$; loosely, the bound one expects for
trying to find a single FP in each of~$r$ independent``rounds'', each round
consisting of~$q$ attempts.

\heading{Construction of PRF with domain separation.}
Let $\Pi_{\mathrm{ds}}= (\calQ,\Rep, \Qry)$ be a set-multiplicity data structure defined on the right-side of Figure~\ref{fig:lin-and-ds}.  As with the preceding theorem, the following says (informally) that if~$F$ is a secure PRF, then this domain-separated PRF construction is correct against adaptive adversaries.  We note that Neidermayer et al. in~\cite{xxx} suggest replacing the linear-PRF hash functions in the previous construction with independently keyed PRFs as a way to thwart their attacks.  The construction we consider now achieves the same end, albeit via explict domain separation rather than separate keys.

\begin{theorem}\label{thm2}
Fix $k,m,n,r>0$, and let $\Pi_{\mathrm{ds}}= (\calQ,\Rep, \Qry)$ be the set-multiplicity data structure in Figure~\ref{fig:lin-and-ds}. Let $\distr{}{}$ be a distribution over~$\mathcal{S}$.  For any adversary~$A$ that makes a total of~$q$ queries to its oracle and has time-complexity~$O(t)$, there exists a prf-adversary ~$B$ (explicitly constructed in the proof of this theorem) such that
\[
\AdvCorrect{\Pi_{\mathrm{ds}},\distr{}{},r}{A} \leq  \AdvPRF{F}{B}  + {\dbinom{q}{r}} \left( (1-e^{-kn/m})^k + O(1/n) \right)^r\,.
\]
Here, $B$ asks $q$ queries, and has time complexity $O(t+qm)$.
\end{theorem}

%%%%%%%%%%
\input{ds-ss-proof}
%%%%%%%%%%
